Toward theories of deliberated preferences

### 2020-08-18 simple.tex
Mostly examples

It is well known that shallow preferences do not always match reflective ones.
Indeed, forming reflective preferences is recognised as a difficult problem in many decision problems, contrary, by definition, to shallow preferences. For example, \citet[p.\ 6]{colell_microeconomic_1995} write, when presenting the preference relation that is the basis for most of their famous book about Microeconomic theory: “Introspection quickly reveals how hard it is to evaluate alternatives that are far from the realm of common experience. It takes work and serious reflection to find out one’s own preferences.” 
It can be expected that some behavior highlighted by psychologists \citep{kahneman_thinking_2012}, such as framing effects, would be rejected by the individuals themselves when pondering their decisions.

@@

\section{Normative theories}
\subsection{Definition}
One of the primitive elements of a normative theory is a set of arguments $\allargs$.
Given a set of arguments $\allargs$, define the set of finite sequences of elements in $\allargs$ (including the empty sequence) as $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$.
Throughout the article, $\N$ includes zero and $\N^* = \N \setminus \set{0}$. Given $j, k \in \N$, the notation $\intvl{j, k}$ represents $\set{l \suchthat j ≤ l ≤ k}$. Note that $\intvl{1, 0} = \emptyset$, and given any set $S$, the notation $S^\emptyset$ represents the empty set. A generic element of $\allhist$ is denoted by $\hist$ and called a path.

@@

@@

An initial argumentation relation ${\gleadsto} \subseteq \allargs × \allprops$ relates some arguments to some propositions. Its role in an empirical theory is to define the propositions that the empirical theory supports (corresponding to $\gPhi$, the range of ${\gleadsto}$) and, for each such proposition, to define the arguments that the theory wants to use as initial arguments to argue in favor of this proposition. The set of initial arguments is $\gleadstoinv(\Phi) = \set{\ar \in \allargs \suchthat \exists \phi \in \Phi \suchthat \ar \gleadsto \phi}$.

@@

Given $i \in I, \phi \in \gPhi, \ar \in \allargs$, define $\di(\ar) \subseteq \powersetz{\alldargs}$ as $\di(\ar) = \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \land \args \ibeats \ar}$. It represents the coalitions of fragments defending $\ar$ (when arguing for $\phi$), from the point of view of $i$, against its effective planned attackers. 
The following proposition shows that these coalitions can be added to $\ar$, and will be useful for other results.
\begin{proposition}
	\label{th:indf}
	$\forall i \in I, \phi \in \gPhi, \ar \in \allargs: \dargs \in \di(\ar) ⇒ \emptyset ≠ \dargs \subseteq \df(\ar)$.
\end{proposition}
\begin{proof}
	I will prove that $\forall \ar \in \allargs: \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar)} = \powersetz{\df(\ar)}$, from which the conclusion follows, and which can be useful to clarify the relationship between $\di(\ar)$ and $\df(\ar)$. Note that by definition of $\df(\ar)$, $\dargs \in \powersetz{\df(\ar)}$ iff $\emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar))$. 
	Thus, we have to prove that
	$\forall \dargs \subseteq \allargs$:
	\begin{equation}
		\exists \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \suchthat \dbeatsinv(\args) = \dargs ⇔ \emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar)).
	\end{equation}

@@

Given $\ar \in \allargs$, define $\dip(\ar) = \set{(\ar + \dargs) \suchthat \dargs \in \di(\ar)}$, representing the subset of defenders of $\ar$ required for convincing $i$ of $\phi$.
Given $i \in I, \phi \in \gPhi$, define $\gargs \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\phi)$ under $\dip$: $\gargs = \cup_{k \in \N} (\dip)^k({\gleadstoinv}(\phi))$.

\begin{definition}[Convincingness]
	An empirical theory is \emph{convincing} iff $\forall i \in I, \phi \in \gPhi, \ar \in \gargs: \ar$ $i$-defends $\phi$.
\end{definition}

A \emph{falsification instance} of an empirical theory is a tuple $(i, \phi, \hist, \ar) \in I × \gPhi × \allhist × \gargs$ such that $\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \phi$. 
I also write that $(i, \phi, \hist, \ar)$ falsifies $\gamma$ to mean that it is a falsification instance of $\gamma$.

\begin{theorem}
	An empirical theory is convincing iff it has no falsification instances.
\end{theorem}
\begin{proof}
	For all $i \in I, \phi \in \gPhi, \ar \in \gargs$: $\ar$ does not $i$-defend $\phi$ iff $\exists \hist \in \allhist \suchthat \left[\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \phi\right]$ iff $\exists \hist \in \allhist$ such that $(i, \phi, \hist, \ar)$ falsifies $\gamma$. Therefore, $\gamma$ is not convincing iff $\exists (i, \phi, \ar) \suchthat \ar$ does not $i$-defend $\phi$ iff $\exists \hist \suchthat (i, \phi, \hist, \ar)$ falsifies $\gamma$ iff $\gamma$ has a falsification instance.
\end{proof}

\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \phi \in \gPhi: [\phi \in \iPhi \land ¬\phi \notin \iPhi]$.
\end{definition}

\begin{lemma}
	\label{th:produce}
	Given $\phi \in \gPhi$, $\ar_0 \in \gargs$ that $i$-defends $\phi$, and $\hist \nileadstost \phi$ with $\ar_0 \in \histend$, define $\args =  \gbeatsinv(\ar_0) \cap \hist$, $\dargs = \dbeatsinv(\args)$, and $\ar_2 = \ar_0 + \dargs$. Then, $\ar_2 \in \gargs \cap \dfp(\ar_0)$.
\end{lemma}
\begin{proof}
	By definition of $\ibeats$, $\args \ibeats \ar_0$.
	Also, because $\ar_0$ $i$-defends $\phi$, $\args ≠ \emptyset$. 
	Therefore, $\dargs \in \di(\ar_0)$.
	It follows that $\dargs \in \df(\ar_0)$, by \cref{th:indf}, and that $\ar_2 \in \dfp(\ar_0)$, by definition of $\dfp$. It also follows that $\ar_2 \in \dip(\ar_0)$, hence, that $\ar_2 \in \gargs$.
\end{proof}

\begin{theorem}
	A normatively adequate and convincing empirical theory is valid.
\end{theorem}
\begin{proof}
	Given any $i \in I$ and $\phi \in \gPhi$, suffices to show that for some $\ar \in \allargs$, $\ar \ileadstost \phi$; the conclusion then follows thanks to simple normative adequacy (that \cref{th:normeqsnorm} shows to be equivalent to normative adequacy).

@@

\subsection{Effectively comparing models}
Given $\phi \in \Phi$ and two empirical theories $\gamma$ and $\delta$ for a given normative theory, say that they disagree about $\phi$ iff $\phi \in \gPhi[\gamma] \land ¬\phi \in \gPhi[\delta]$. Say that they are incompatible iff there is at least one $\phi \in \Phi$ that they disagree about.

@@

Define $\Phi_\allalts = \set{\phi_{B, a}, \phi_{B, ¬a} \suchthat \emptyset ≠ B \subseteq \allalts, a \in B}$.

@@

The observation protocol goes as follows. Given $i \in I$, $\emptyset ≠ B \subseteq \allalts$ and $\hist \in \allhist$, the individual $i$, after having been presented the arguments in the sequence $\hist$, is asked to choose among a subset of elements from $B$ among which she is indifferent. Define $\emptyset ≠ c^\hist_i(B) \subseteq B$ as the subset that $i$ chooses after exposure to $\hist$.
Define $\hist \ileadsto \phi_{B, a}$ iff $a \in c^\hist_i(B)$.
Define $\hist \ileadsto \phi_{B, ¬a}$ iff $a \notin c^\hist_i(B)$.

\begin{remark}
	This does not work, because $\gamma$ supporting $\phi_{\set{a, b, c, d}, ¬b}$, $\phi_{\set{a, b, c, d}, ¬c}$ and $\phi_{\set{a, b, c, d}, ¬d}$ may be valid even if $i$ never chooses $a$: $i$ may satisfy $¬b$ (by picking $c$) when given an argument for $¬b$, satisfy both $¬b$ and $¬c$ (by picking $d$) when given arguments for $¬b$ and $¬c$, and so on. The individual $i$ is not obliged to satisfy the three arguments for $¬b$, $¬c$ and $¬d$ simultaneously, as these arguments are required to count only when they are at the end of the sequence.
\end{remark}

This defines normative theories about $\allalts$. Any empirical theory for a normative theory about $\allalts$ is said to be an empirical theory about $\allalts$.

Given an empirical theory about $\allalts$ and $a, b \in \allalts$, define $a \gpst b$ iff $\forall \set{a, b} \subseteq B: \phi_{B, ¬b} \in \gPhi$.
Define $a \gind b$ iff either $a = b$ or $\forall \set{a, b} \subseteq B: [\phi_{B, a} \in \gPhi \land \phi_{B, b} \in \gPhi] \lor [\exists c \in B \suchthat c \gpst a \land c \gpst b]$.
Define ${\gpeq} = {\gpst} \cup {\gind}$.

An empirical theory about $\allalts$ is said to be coherent iff $\forall \emptyset ≠ B \subseteq \allalts$: $¬(\forall a \in B: \phi_{B, ¬a} \in \gPhi) \land ¬(\exists a \in B \suchthat \phi_{B, a} \in \gPhi \land \phi_{B, ¬a} \in \gPhi)$. Coherence is required for validity. Note that it is not an empirical property, in the sense that it can be verified without experiments, as it is independent of $\ileadsto$.

@@

Given $i \in I$ and a bundle $B$, say that a subset $B' \subseteq B$ is a deliberated restricted choice set for $i$ given $B$ iff $[\forall a \in B \setminus B': \phi_{B, ¬a} \in \iPhi] \lor [\exists a \in B' \suchthat \phi_{B, a} \in \iPhi]$. Restricting consideration to deliberated restricted choice sets respects $i$’s \ac{DJ} in the sense that such sets exclude from $B$ only alternatives that $i$ deliberately excludes from possible choice among the bundle $B$, or include an alternative that $i$ deliberately chooses when choosing from the bundle $B$.

@@

\begin{theorem}[Deliberated preference]
	\label{th:dp}
	Let $\gamma$ be a valid empirical theory for a normatively adequate normative theory about $\allalts$, and such that $\gpeq$ is transitive. Let $r$ be a choice function such that, given any $B \in \powersetz[\mathit{F}]{\allalts}$, $r(B)$ contains exactly one arbitrary argument of each undominated equivalence classes of $B$, where $\gind$ determines the equivalence classes and $\gpst$ determines the undominated elements. Then, $\forall B \in \powersetz[\mathit{F}]{\allalts}$, $r(B)$ is a deliberated restricted choice set for $i$ given $B$.
\end{theorem}
\begin{proof}
	Observe that under these hypothesis, $\gpeq$ is a transitive, reflexive binary relation on $\allalts$, $\gpst$ is its asymmetric part and $\gind$ is its symmetric part. That $\gpst$ is asymmetric follows from coherence, which requires, considering $B = \set{a, b}$, that $\phi_{B, ¬a} \notin \gPhi$ or $\phi_{B, ¬b} \notin \gPhi$ (hence $b \ngpst a$ or $a \ngpst b$). Coherence itself follows from validity.

@@

	Given any bundle $B \in \powersetz[\mathit{F}]{\allalts}$, we now have to show that $B' = r(B)$ satisfies $[\forall a \in B \setminus B': \phi_{B, ¬a} \in \iPhi] \lor [\exists a \in B' \suchthat \phi_{B, a} \in \iPhi]$. Because $\gamma$ is valid, suffices to show that $[\forall a \in B \setminus B': \phi_{B, ¬a} \in \gPhi] \lor [\exists a \in B' \suchthat \phi_{B, a} \in \gPhi]$.
	Consider any $a \in B \setminus B'$. Because $B'$ is finite, for some $b \in B'$, $b \gpst a$ or $b \gind a$.

	If $b \gpst a$, and because $\set{a, b} \in B$, $\phi_{B, ¬a} \in \gPhi$.

	If $b \gind a$, and because $\set{a, b} \in B$, $\phi_{B, b} \in \gPhi$ (because $\nexists c \in B \suchthat c \gpst b$, otherwise $b \notin B'$).
\end{proof}
The phrasing of this theorem considers only finite bundles for simplicity. It more generally applies to bundles that are closed according to $\gpst$. \TODO{Phrase this correctly.}

\begin{proposition}[Completeness implies transitivity]
	\label{th:comptrans}
	If $\gpeq$ is complete ($\forall a, b \in \allalts: a \gpeq b \lor b \gpeq a$) and $\gamma$ is coherent, then it is transitive.
\end{proposition}
\begin{proof}
	Consider any $a, b, c \in \allalts$ with $a \gpeq b \gpeq c$ and let us show that $a \gpeq c$.

	Observe first that given any bundle $B \subseteq \allalts$ with $x, y \in B$, when $x \gpeq y$, $\phi_{B, ¬y} \notin \gPhi ⇒ \phi_{B, ¬x} \notin \gPhi$. 
	Indeed, when $x ≠ y$,
	because $x \gpeq y$, either $\phi_{B, ¬y} \in \gPhi$ or $\phi_{B, x} \in \gPhi$; thus, if $\phi_{B, ¬y} \notin \gPhi$, then $\phi_{B, x} \in \gPhi$, hence, by coherence, $\phi_{B, ¬x} \notin \gPhi$.

	Consider now the bundle $B = \set{a, b, c}$. By coherence, $[\phi_{B, ¬a} \notin \gPhi \lor \phi_{B, ¬b} \notin \gPhi \lor \phi_{B, ¬c} \notin \gPhi]$. The above observation leads to $\phi_{B, ¬c} \notin \gPhi ⇒ \phi_{B, ¬b} \notin \gPhi$ and $\phi_{B, ¬b} \notin \gPhi ⇒ \phi_{B, ¬a} \notin \gPhi$. Therefore, $\phi_{B, ¬a} \notin \gPhi$.

	Completeness implies that $a \gpeq c \lor c \gpeq a$, or equivalently, $a \gpeq c \lor c \gpst a \lor c \gind a$. Excluding $c \gpst a$ thanks to $\phi_{B, ¬a} \notin \gPhi$, we obtain $a \gpeq c$.
\end{proof}

@@

When $\gpeq$ is not transitive, it still permits to define choice functions that could be said to respect, in some reasonable sense, the deliberated preference of the individuals. However, this claim is more arguable, as it is generally (though by all means not unanimously) considered in the literature that a reflective preference should be transitive. Further explorations of this possibility are left for future articles.

The claims of an empirical theory about $\allalts$ can sometimes be defined on the basis of a preference ordering $\peq$. In such a case, the claims …

\subsection{Satisfaction}
Consider a normative theory of preferences about a given set of alternatives $\allalts$. Let $\gamma$ be the following empirical theory for that normative theory.

Define $C$ a set of criteria, functions $c: \allalts → \R$, each function measuring the performances of the alternatives according to some point of view. Let $t \in \R^C$ be a set of thresholds, thus, consisting of one real valued threshold $t_c \in \R$ per criterion. Say that $a \in \allalts$ is satisfying iff $\forall c \in C: c(a) ≥ t_c$.

The theory $\gamma$ claims, intuitively, that individuals $i \in I$ are satisficers using some defined criteria $C$ and thresholds $t = (t_c)_{c \in C}$. The set $\gPhi$ is defined as $\set{\phi_{a > b} \suchthat a \text{ is satisficing and $b$ is not}}$. Given $a, b \in \allalts$, with $a$ satisficing and $b$ not, define $\ar_{a > b}$ as the text: “$a$ has performance $c_1(a)$ on criterion $c_1$, and it is reasonable to consider that having at least $t_{c_1}$ is satisfying enough from the point of view captured by $c_1$, (… same phrase for the other criteria…), therefore, it is a fully satisfactory alternative; whereas $b$ is not satisfactory on criteria (… enumeration of the criteria such that $c(b) < t_{c}$).” 
Define ${\gleadsto} = \set{\ar_{a > b}, \phi_{a > b}) \suchthat a \text{ is satisficing and $b$ is not}}$.
Define $\alldargs = \gbeats = \dbeats = {+} = \emptyset$: the theory does not consider any counter-argument as potentially damaging. 
This defines the theory $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$.

Observe that $\gamma$ does not claim a complete preference, but, whenever there is exactly one satisficing alternative in $\allalts$, it still determines a deliberated preferred choice for $i$ among $\allalts$.

>>>>

It is well known that shallow preferences do not always match reflective ones.
Indeed, forming reflective preferences is recognised as a difficult problem in many decision problems, contrary, by definition, to shallow preferences. For example, \citet[p.\ 6]{colell_microeconomic_1995} write, when presenting the preference relation that is the basis for most of their famous book about Microeconomic theory: “Introspection quickly reveals how hard it is to evaluate alternatives that are far from the realm of common experience. It takes work and serious reflection to find out one’s own preferences.” 
In the words of \citet[p.\ 9]{bell_descriptive_1988}, “Our decision makers are not economic automatons; they make mistakes, have remorse, suffer anxieties, and cannot make up their minds” (cited by \citet[p.\ 87]{deparis_etude_2012}).
Indeed, it is to be expected that some behavior highlighted by psychologists \citep{kahneman_thinking_2012}, such as framing effects, would be rejected by the individuals themselves when pondering their decisions.

@@

\section{Normative theories}
\subsection{Definition}
\TODO{Consider using only $\phi$ being sure, and dropping everything related to possibility.}

One of the primitive elements of a normative theory is a set of arguments $\allargs$.
Given a set of arguments $\allargs$, define the set of finite sequences of elements in $\allargs$ (including the empty sequence) as $\allhist = \bigcup_{k \in \N} \allargs^{\intvl{1, k}}$.
Throughout the article, $\N$ includes zero and $\N^* = \N \setminus \set{0}$. Given $j, k \in \N$, the notation $\intvl{j, k}$ represents $\set{l \suchthat j ≤ l ≤ k}$. Note that $\intvl{1, 0} = \emptyset$, and given any set $S$, the notation $S^\emptyset$ represents the empty set. A generic element of $\allhist$ is denoted by $\hist$ and called a path.

@@

\begin{remark}
	I’d rather drop the negation and observe that if the protocol is defined so that $\hist \ileadsto \phi_1 ⇒ \hist \nileadsto \phi_2$, then $\phi_1 \in \iPhi ⇒ \phi_2 \notin \iPhi$. Observe that this is always possible. The definitions about completeness should use a negation operator in supplement to the normative theory tuple.
\end{remark}

@@

An initial argumentation relation ${\gleadsto} \subseteq \allargs × \allprops$ relates some arguments to some propositions. Its role in an empirical theory is to define the propositions that the empirical theory supports (corresponding to $\gPhie$, the range of ${\gleadsto}$) and, for each such proposition, to define the arguments that the theory wants to use as initial arguments to argue in favor of this proposition. The set of initial arguments is $\gleadstoinv(\Phi) = \set{\ar \in \allargs \suchthat \exists \phi \in \Phi \suchthat \ar \gleadsto \phi}$.

@@

Given $i \in I, \phi \in \gPhie, \ar \in \allargs$, define $\di(\ar) \subseteq \powersetz{\alldargs}$ as $\di(\ar) = \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \land \args \ibeats \ar}$. It represents the coalitions of fragments defending $\ar$ (when arguing for $\phi$), from the point of view of $i$, against its effective planned attackers. 
The following proposition shows that these coalitions can be added to $\ar$, and will be useful for other results.
\begin{proposition}
	\label{th:indf}
	$\forall i \in I, \phi \in \gPhie, \ar \in \allargs: \dargs \in \di(\ar) ⇒ \emptyset ≠ \dargs \subseteq \df(\ar)$.
\end{proposition}
\begin{proof}
	I will prove that $\forall \ar \in \allargs: \set{\dbeatsinv(\args) \suchthat \emptyset ≠ \args \subseteq \gbeatsinv(\ar)} = \powersetz{\df(\ar)}$, from which the conclusion follows, and which can be useful to clarify the relationship between $\di(\ar)$ and $\df(\ar)$. Note that by definition of $\df(\ar)$, $\dargs \in \powersetz{\df(\ar)}$ iff $\emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar))$. 
	Thus, we have to prove that
	$\forall \dargs \subseteq \allargs$:
	\begin{equation}
		\exists \emptyset ≠ \args \subseteq \gbeatsinv(\ar) \suchthat \dbeatsinv(\args) = \dargs ⇔ \emptyset ≠ \dargs \subseteq \dbeatsinv(\gbeatsinv(\ar)).
	\end{equation}

@@

Given $\ar \in \allargs$, define $\dip(\ar) = \set{(\ar + \dargs) \suchthat \dargs \in \di(\ar)}$, representing the subset of defenders of $\ar$ required for convincing $i$ of $\phi$.
Given $i \in I, \phi \in \gPhie$, define $\gargs \subseteq \allargs$ as the transitive closure of ${\gleadstoinv}(\phi)$ under $\dip$: $\gargs = \cup_{k \in \N} (\dip)^k({\gleadstoinv}(\phi))$.

\begin{definition}[Convincingness]
	An empirical theory is \emph{convincing} iff $\forall i \in I, \phi \in \gPhie, \ar \in \gargs: \ar$ $i$-defends $\phi$.
\end{definition}

A \emph{falsification instance} of an empirical theory is a tuple $(i, \phi, \hist, \ar) \in I × \gPhie × \allhist × \gargs$ such that $\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \phi$. 
I also write that $(i, \phi, \hist, \ar)$ falsifies $\gamma$ to mean that it is a falsification instance of $\gamma$.

\begin{theorem}
	An empirical theory is convincing iff it has no falsification instances.
\end{theorem}
\begin{proof}
	For all $i \in I, \phi \in \gPhie, \ar \in \gargs$: $\ar$ does not $i$-defend $\phi$ iff $\exists \hist \in \allhist \suchthat \left[\gbeatsinv(\ar) \cap \hist = \emptyset \land \ar \in \histend \land \hist \nileadstost \phi\right]$ iff $\exists \hist \in \allhist$ such that $(i, \phi, \hist, \ar)$ falsifies $\gamma$. Therefore, $\gamma$ is not convincing iff $\exists (i, \phi, \ar) \suchthat \ar$ does not $i$-defend $\phi$ iff $\exists \hist \suchthat (i, \phi, \hist, \ar)$ falsifies $\gamma$ iff $\gamma$ has a falsification instance.
\end{proof}

\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \phi \in \gPhie: [\phi \in \iPhi \land ¬\phi \notin \iPhi]$.
\end{definition}

\begin{lemma}
	\label{th:produce}
	Given $\phi \in \gPhie$, $\ar_0 \in \gargs$ that $i$-defends $\phi$, and $\hist \nileadstost \phi$ with $\ar_0 \in \histend$, define $\args =  \gbeatsinv(\ar_0) \cap \hist$, $\dargs = \dbeatsinv(\args)$, and $\ar_2 = \ar_0 + \dargs$. Then, $\ar_2 \in \gargs \cap \dfp(\ar_0)$.
\end{lemma}
\begin{proof}
	By definition of $\ibeats$, $\args \ibeats \ar_0$.
	Also, because $\ar_0$ $i$-defends $\phi$, $\args ≠ \emptyset$. 
	Therefore, $\dargs \in \di(\ar_0)$.
	It follows that $\dargs \in \df(\ar_0)$, by \cref{th:indf}, and that $\ar_2 \in \dfp(\ar_0)$, by definition of $\dfp$. It also follows that $\ar_2 \in \dip(\ar_0)$, hence, that $\ar_2 \in \gargs$.
\end{proof}

\begin{theorem}
	A normatively adequate and convincing empirical theory is valid.
\end{theorem}
\begin{proof}
	Given any $i \in I$ and $\phi \in \gPhie$, suffices to show that for some $\ar \in \allargs$, $\ar \ileadstost \phi$; the conclusion then follows thanks to simple normative adequacy (that \cref{th:normeqsnorm} shows to be equivalent to normative adequacy).

@@

\subsection{Effectively comparing models}
Given $\phi \in \Phi$ and two empirical theories $\gamma$ and $\delta$ for a given normative theory, say that they disagree about $\phi$ iff $\phi \in \gPhie[\gamma] \land ¬\phi \in \gPhie[\delta]$. Say that they are incompatible iff there is at least one $\phi \in \Phi$ that they disagree about.

@@

Define $\Phi_\allalts = \set{\phi_{B, \choices} \suchthat \emptyset ≠ B \subseteq \allalts, \choices \subseteq \powersetz{B}}$.

@@

The observation protocol goes as follows. Given $i \in I$, $\emptyset ≠ B \subseteq \allalts$ and $\hist \in \allhist$, the individual $i$, after having been presented the arguments in the sequence $\hist$, is asked to choose among a subset of elements from $B$ among which she is indifferent. Define $\emptyset ≠ c^\hist_i(B) \subseteq B$ as the subset that $i$ chooses after exposure to $\hist$.
Define $\hist \ileadsto \phi_{B, \choices}$ iff $c^\hist_i(B) \in \choices$.

This defines normative theories about $\allalts$. Any empirical theory for a normative theory about $\allalts$ that satisfies $\forall \emptyset ≠ B \subseteq \allalts: \exists ! \emptyset ≠ \choices \subseteq \powersetz{B} \suchthat \phi_{B, \choices} \in \gPhie$ is said to be an empirical theory about $\allalts$. The notation $\exists!$ means: there exists a unique element. The requirement that there is at least one is innocuous, as choosing $\choices = \powersetz{B}$ ensures that $\phi_{B, \choices} \in \iPhi$, by definition of $\ileadsto$.

An empirical theory about $\allalts$ thus associates to each possible bundle $B$ a set of possible choices $\choices$, whose semantics is that the theory claims that the deliberated choice of $i$, given the bundle $B$, lies among $\choices$. Let $\gPhi(B) \subseteq \powersetz{B}$ denote this set of possible choices, thus $\gPhi(B) = \choices  ⇔ \phi_{B, \choices} \in \gPhie$.

Given an empirical theory about $\allalts$ and $a, b \in \allalts$, define $a \gpst b$ iff $\forall \set{a, b} \subseteq B: b \notin \bigcup \gPhi(B)$. Thus, $a \gpst b$ iff $\gamma$ claims that $b$ is not deliberately chosen in presence of $a$.
Define $a \gind b$ iff $\forall \set{a, b} \subseteq B: [\forall C \in \choices: a \in C ⇔ b \in C]$.
Define ${\gpeq} = {\gpst} \cup {\gind}$.

@@

Given $i \in I$ and a bundle $B$, say that a subset $B' \subseteq B$ is a deliberated restricted choice set for $i$ given $B$ iff 
$\exists \phi_{B, \choices} \in \iPhi \suchthat \forall C \in \choices: C \cap B' ≠ \emptyset$.
Restricting consideration to deliberated restricted choice sets respects $i$’s \ac{DJ} in the sense that it necessarily includes at least one alternative that is a deliberated choice of $i$.

@@

\begin{theorem}[Deliberated preference]
	\label{th:dp}
	Let $\gamma$ be a valid empirical theory for a normatively adequate normative theory about $\allalts$, and such that $\gpeq$ is transitive. Let $r$ be a choice function such that, given any $B \in \powersetz[\mathit{F}]{\allalts}$, $r(B)$ contains exactly one arbitrary argument of each undominated equivalence classes of $B$, where $\gind$ determines the equivalence classes and $\gpst$ determines the undominated elements. Then, $\forall B \in \powersetz[\mathit{F}]{\allalts}$, $r(B)$ is a deliberated restricted choice set for $i$ given $B$.
\end{theorem}
\begin{proof}
	Observe that under these hypothesis, $\gpeq$ is a transitive, reflexive binary relation on $\allalts$, $\gpst$ is its asymmetric part and $\gind$ is its symmetric part. 
	$\gpeq$ is reflexive because $\gind$ is.
	That $\gpst$ is asymmetric is seen by considering, given any $a, b \in \allalts$, the bundle $B = \set{a, b}$: pick any $C \in \gPhi(B) \subseteq \powersetz{B}$, thus, $C \in \powersetz{B}$, hence, $a \in C$ or $b \in C$, and therefore, $b \ngpst a$ or $a \ngpst b$.

@@

	Given any bundle $B \in \powersetz[\mathit{F}]{\allalts}$, we now have to show that $B' = r(B)$ satisfies 
	$\exists \phi_{B, \choices} \in \iPhi \suchthat \forall C \in \choices: C \cap B' ≠ \emptyset$.
%	$[\forall a \in B \setminus B': \phi_{B, ¬a} \in \iPhi] \lor [\exists a \in B' \suchthat \phi_{B, a} \in \iPhi]$. 
	Because $\gamma$ is valid, suffices to show that, given $\choices = \gPhi(B) \subseteq \powersetz{B}$, 
	$\forall C \in \choices: C \cap B' ≠ \emptyset$.
%	$[\forall a \in B \setminus B': \phi_{B, ¬a} \in \gPhie] \lor [\exists a \in B' \suchthat \phi_{B, a} \in \gPhie]$.
	Pick any $C \in \choices$ and any $a \in C$ (thus, $a \in B$). By definition of $\gpst$, $\nexists b \in B \suchthat b \gpst a$ as otherwise, $a \notin C$.
	Note that, by construction of $B'$, as $a \in B$ is undominated, it belongs to some equivalence class of which a member has been brought in $B'$. In other words, $\exists b \in B' \suchthat b \gind a$. Therefore, by definition of $\gind$, $b \in C$. Thus, $b \in C \cap B'$.
\end{proof}
The phrasing of this theorem considers only finite bundles for simplicity. It more generally applies to bundles that are closed according to $\gpst$. \TODO{Phrase this correctly.}

\begin{proposition}[Completeness implies transitivity]
	\label{th:comptrans}
	If $\gpeq$ is complete ($\forall a, b \in \allalts: a \gpeq b \lor b \gpeq a$) and $\gamma$ is coherent, then it is transitive.
\end{proposition}
\begin{proof}
	Consider any $a, b, c \in \allalts$ with $a \gpeq b \gpeq c$ and let us show that $a \gpeq c$.
	Consider the bundle $B = \set{a, b, c}$ and let $\choices = \gPhi(B)$.

	Observe first that if $x, y \in B$ and $x \gpeq y$, $y \in \bigcup \choices ⇒ x \in \bigcup \choices$. 
	Indeed, if $x \gpst y$, $y \notin \bigcup \choices$, and if $x \gind y$, $y \in \bigcup \choices ⇔ x \in \bigcup \choices$ as $x$ and $y$ can then be treated interchangeably considering their memberships to elements of $\choices$.

	Because $\bigcup \choices ≠ \emptyset$, $[a \in \bigcup \choices \lor b \in \bigcup \choices \lor c \in \bigcup \choices]$. The above observation leads to $c \in \bigcup \choices ⇒ b \in \bigcup \choices$ and $b \in \bigcup \choices ⇒ a \in \bigcup \choices$. Therefore, $a \in \bigcup \choices$.

	Completeness implies that $a \gpeq c \lor c \gpeq a$, or equivalently, $a \gpeq c \lor c \gpst a \lor c \gind a$. Excluding $c \gpst a$ thanks to $a \in \bigcup \choices$, we obtain $a \gpeq c$.
\end{proof}

@@

Note that $\gpeq$ is independent of $\ileadsto$; thus, any property of $\gpeq$, such as transitiveness or completeness, is a priori (as opposed to empirical), meaning that its satisfaction can be verified without experiments.

When $\gpeq$ is not transitive, choice functions could still be defined that could be said to respect, in some reasonable sense, the deliberated preference of the individuals. However, this claim is more arguable, as it is generally (though by all means not unanimously) considered in the literature that a reflective preference should be transitive. Further explorations of this possibility are left for future articles.

In what follows, the claims of an empirical theory about $\allalts$ will be defined on the basis of a transitive preference ordering $\peq$ on $\allalts$, as follows. Given $\emptyset ≠ B \subseteq \allalts$, define $\gPhi(B) = \set{\cup \choices' \suchthat \emptyset ≠ \choices' \subseteq \choices}$, where $\choices$ are the undominated equivalence classes in $B$ according to $\peq$ (thus $\choices = \max_{\pst^*}(B / {\ind})$, using the notations defined in \cref{th:dp}). The check that ${\peq} = {\gpeq}$ is left to the reader. Thus, all the examples that follow have a transitive $\gpeq$ and therefore claim to represent a deliberated preference (when valid) in the sense defined in this section.

\subsection{A theory based on Satisfaction}
Consider a normative theory of preferences about a given set of alternatives $\allalts$. Let $\gamma$ be the following empirical theory for that normative theory.

Define $C$ a set of criteria, functions $c: \allalts → \R$, each function measuring the performances of the alternatives according to some point of view. Let $t \in \R^C$ be a set of thresholds, thus, consisting of one real valued threshold $t_c \in \R$ per criterion. Say that $a \in \allalts$ is satisfying iff $\forall c \in C: c(a) ≥ t_c$.
 Let $A$ denote the satisfying alternatives.

The theory $\gamma$ claims, intuitively, that individuals $i \in I$ are satisficers using some defined criteria $C$ and thresholds $t = (t_c)_{c \in C}$. Define $\peq$ as $a \peq b$ iff $a \in A \lor b \notin A$. Note that $\peq$ is complete and contains at most two equivalence classes: $A$, and $\allalts \setminus A$. This defines the set of claims that $\gamma$ supports, namely, for any $\emptyset ≠ B \subseteq \allalts$, $\gPhi(B) = B \cap A$.

Given $a \in A$, define the fragment $\ar_a$ as the text: “$a$ has performance $c_1(a)$ on criterion $c_1$, and it is reasonable to consider that having at least $t_{c_1}$ is satisfying enough from the point of view captured by $c_1$, [… same phrase for the other criteria…], therefore, it is a fully satisfactory alternative.”
Given $b \in \allalts \setminus A$, define $\ar_b$ as the text: “$b$ is not satisfactory on the criterion [some criterion such that $c(b) < t_{c}$]: the performance $t_c$ should be attained for considering $b$ as satisfying on that point of view, but it obtains only $c(b)$ on that criterion [… same phrase for the other criteria such that $c(b) < t_c$].”

Given a bundle $B$, define $\ar_B$ as the text obtained by concatenating all fragments in $\set{\ar_a \suchthat a \in B}$.

%Given $a \in A, b \in \allalts \setminus A$, define $\ar_{a > b}$ as the text: “$a$ has performance $c_1(a)$ on criterion $c_1$, and it is reasonable to consider that having at least $t_{c_1}$ is satisfying enough from the point of view captured by $c_1$, [… same phrase for the other criteria…], therefore, it is a fully satisfactory alternative; whereas $b$ is not satisfactory on criteria [… enumeration of the criteria such that $c(b) < t_{c}$].” 
Define ${\gleadsto} = \set{(\ar_B, \gPhi(B)) \suchthat \emptyset ≠ B \subseteq \allalts}$.
Define $\alldargs = {\gbeats} = {\dbeats} = {+} = \emptyset$: the theory does not consider any counter-argument as potentially damaging. 
This defines an empirical theory $(\gleadsto, \gbeats, \alldargs, \dbeats, +)$ about $\allargs$.
